{\fontfamily{ptm}\selectfont Linear Convergence of Stochastic Nesterov's Accelerated Proximal Gradient method under Interpolation Hypothesis}

### 2025-07-08 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, changes

        \begin{definition}[Proximal gradient operator]\label{def:pg-opt}
            Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}. 
            Let $\beta > 0$. 
            Then, we define the proximal gradient operator $T_{\beta}$ as 
            \begin{align*}
                T_\beta (x | F) &= \argmin{z} \left\lbrace
                    g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{\beta}{2}\Vert z - x\Vert^2
                \right\rbrace. 
            \end{align*}
        \end{definition}
        \begin{remark}
            If the function $g \equiv 0$, then it yields the gradient descent operator $T_\beta(x) = x - \beta^{-1}\nabla f(x)$. 
            In the context where it's clear what the function $F = f + g$ is, we simply write $T_\beta(x)$ for short. 
        \end{remark}

@@

        \begin{theorem}[proximal gradient inequality]\label{thm:pg-ineq}
            Let function $F$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}, so it's $\mu \ge 0$ strongly convex. 
            For all $x\in \RR^n$, define $x^+ = T_L(x)$, then there exists a $B \ge 0$ such that $D_f(x^+, x) \le B/2 \Vert x^+ - x\Vert^2$. 
            Then, for all $z \in \RR^n$ it satisfies proximal gradient inequality at point $x$:  
            \begin{align*}
                0&\le F(z) - F(x^+) - \frac{B}{2}\Vert z - x^+\Vert^2  + \frac{B - \mu}{2}\Vert z - x\Vert^2
                \\
                &=  F(z) - F(x^+) - \langle B(x - x^+), z - x\rangle
                - \frac{\mu}{2}\Vert z - x\Vert^2
                - \frac{B}{2}\Vert x - x^+\Vert^2. 
            \end{align*}
            Since $f$ is assumed to be $L$ Lipschitz smooth, the above condition is true for all $x, y \in \RR^n$ for all $B \ge L$. 
        \end{theorem}
        \begin{remark}
            The theorem is the same as in Nesterov's book \cite[Theorem 2.2.13]{nesterov_lectures_2018}, but with the use of proximal gradient mapping and proximal gradient instead of project gradient hence making it equivalent to the theorem in Beck's book \cite[Theorem 10.16]{beck_first-order_2017}. 
            The only generalization here is parameter $B$ which made to accommodate algorithm that implements Definition \ref{def:snapg-v2} with line search routine to determine $L_k$. 
            Each of the reference books gives a proof of the theorem. 
            But for the best consistency in notations, see Theorem 2.3 in Li and Wang \cite{li_relaxed_2025}. 
        \end{remark}
        \begin{theorem}[Jensen's inequality]\label{thm:jesen}
            Let $F: \RR^n \rightarrow \overline \RR$ be a $\mu \ge 0$ strongly convex function. 
            Then, it is equivalent to the following condition. 
            For all $x, y \in \RR^n$, $\lambda \in (0, 1)$ it satisfies the inequality 
            \begin{align*}
                (\forall \lambda \in [0, 1])\; 
                F(\lambda x + (1 - \lambda)y) \le \lambda F(x) + (1 - \lambda)F(y) -\frac{\mu\lambda(1 - \lambda)}{2} \Vert y - x\Vert^2. 
            \end{align*}
        \end{theorem}
        \begin{remark}
            If $x, y$ is out of $\dom F$, the inequality still work by convexity. 
        \end{remark}

        \begin{definition}[semi quasi strong convexity (\textcolor{red}{NEW})]\label{def:sq-scnvx}
            A function $f:\RR^n \rightarrow \overline \RR$ is semi quasi strongly convex if, there exists a constant $\mu \ge 0$ and, a non-zero linear mapping $A: \RR^n\rightarrow \RR^m$ such that for all $x \in \RR^n, y \in RR^n$ and $\lambda \in [0, 1]$ it satisfies:

@@

                F(\lambda x + (1 - \lambda)y) \le \lambda F(x) + (1 - \lambda)F(y) -\frac{\mu\lambda(1 - \lambda)}{2} \Vert A(y - x)\Vert^2. 

@@

            It is strong convexity but instead of being strongly convex to the function $\mu/2\Vert \cdot\Vert^2$, it's $x \mapsto \mu/2\Vert Ax \Vert^2$ instead. 
            This is an example of uniform convexity because it also fits in Bauchke \cite[Definition 10.7]{bauschke_convex_2017}, with $\phi = \Vert A\Vert^2\Vert x - y\Vert^2$. 

@@

@@

                \item ... \todoinline{This part is not finished yet. }

@@

        \end{remark}
        % --------------------------------------------------------------------------------------------------------------
        \begin{definition}[SNAPG-V2]\label{def:snapg-v2}
            Let $F$ satisfies Assumption \ref{ass:sum-of-many}. 
            Let $(I_k)_{k \ge 0}$ be a list of i.i.d random variables uniformly sampled from set $\{0, 1, 2, \cdots, n\}$. 
            Initialize $v_{-1} = x_{-1}, \alpha_0 = 1$. 
            \textcolor{red}{Let $\tilde \mu \ge 0$ be a fixed constant.}
            The SNAPG generates the sequence $(y_k, x_k, v_k)_{k \ge 0}$ such that for all $k \ge 0$ they satisfy: 
            \begin{align*}
                & \alpha_k \in (0, 1): (L_{k - 1}/L_k)(1 - \alpha_{k})\alpha_{k - 1}^2 = \alpha_{k}\left(\alpha_{k} - \tilde \mu/L_k\right), \\
                & \tau_k = L_k(1 - \alpha_k)\left(L_k \alpha_k - \mu^{(I_k)}\right)^{-1}, \\
                & y_k = (1 + \tau_k)^{-1}v_{k - 1} + \tau_k(1 + \tau_k)^{-1}x_{k - 1}, \\
                & L_k > 0: D_f(x_k, y_k) \le L_k/2\Vert y_k - x_k\Vert^2, \\
                & x_k =  T_{L_k}(y_k | F_{I_k}), \\
                & v_k = x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
            \end{align*}
        \end{definition}
        \begin{remark}
            $\tilde \mu_k, L_k$ are not necessary a random variable because they are determined by a line-search like conditions, consequently $(\alpha_k)_{k\ge 0}$, whether they are a random variable depends on the line search procedures. 
            Otherwise, all the iterates $(x_k, y_k, z_k)$ are random variable determined by $I_k$ when conditioned on all previous $I_{k - 1}, I_{k - 2}, \ldots, I_{0}$. 
            \par
            \textcolor{red}{NEW}. One may notice that $\alpha_k$ requires $L_k$ which comes before $L_k, x_k$ which are needed in advanced for $\alpha_k$. 
            This may seems off since no algorithm can know what $L_k$ to choose in advanced to determine the line search. 
            But, it is important to note that in here, we defined a sequence of conditions on the iterates $x_k, y_k, z_k$, and auxilary sequences $\alpha_k, L_k$ which is not a definition of any algorithm. 
            It is quantifying the conditions needed for an algorithm that actually implements it.
            \par
            For the trivial case where we don't need to worry about it is when $L_k = \max_{i = 1, \ldots, n} K^{(i)}$. 
            See Chambolle, Calatroni \cite{calatroni_backtracking_2019} for an implementations of linear search with backtracking for the FISTA algorithm, it is how one would implement it in the deterministic case. 
        \end{remark}

@@

>>>>

        \begin{definition}[Proximal gradient operator]\label{def:pg-opt}
            Suppose $F = f + g$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}. 
            Let $\beta > 0$. 
            Then, we define the proximal gradient operator $T_{\beta}$ as 
            \begin{align*}
                T_\beta (x | F) &= \argmin_{z} \left\lbrace
                    g(z) + f(x) + \langle \nabla f(x), z - x\rangle + \frac{\beta}{2}\Vert z - x\Vert^2
                \right\rbrace. 
            \end{align*}
        \end{definition}
        \begin{remark}
            If the function $g \equiv 0$, then it yields the gradient descent operator $T_\beta(x) = x - \beta^{-1}\nabla f(x)$. 
            In the context where it's clear what the function $F = f + g$ is, we simply write $T_\beta(x)$ for short. 
        \end{remark}

@@

        \begin{theorem}[Jensen's inequality]\label{thm:jesen}
            Let $F: \RR^n \rightarrow \overline \RR$ be a $\mu \ge 0$ strongly convex function. 
            Then, it is equivalent to the following condition. 
            For all $x, y \in \RR^n$, $\lambda \in (0, 1)$ it satisfies the inequality 

@@

                (\forall \lambda \in [0, 1])\; 
                F(\lambda x + (1 - \lambda)y) \le \lambda F(x) + (1 - \lambda)F(y) -\frac{\mu\lambda(1 - \lambda)}{2} \Vert y - x\Vert^2. 
            If $x, y$ is out of $\dom F$, the inequality still work by convexity. 
        \begin{definition}[semi quasi strong convexity Jensen inequality \textcolor{red}{NEW}]\label{def:sq-scnvx}
            Let $m, n \in \N$. 
            A function $f:\RR^n \rightarrow \overline \RR$ is semi quasi strongly convex with linear mapping $A:\RR^m \rightarrow \RR^n$ if for all $x \in \RR^n, y \in \RR^n$ and $\lambda \in [0, 1]$ it satisfies:
                F(\lambda x + (1 - \lambda)y) \le \lambda F(x) + (1 - \lambda)F(y) -\frac{\lambda(1 - \lambda)}{2} \Vert Ay - Ax\Vert^2. 

@@

            Choose $A = \sqrt{\mu} I$, where $I$ is the identity matrix then it is equivalent to $\mu$ strongly convex. 
            This is an example of uniform convexity because there exists some $\sigma \ge 0$ such that it fits in Bauschke \cite[Definition 10.7]{bauschke_convex_2017} with $\phi(z) = \sigma^2z^2$. 
            It is called ``Semi'' because of the semi norm induced by matrix $A$, followed by ``Quasi strongly convex'' because the quasi norm $\Vert Ax\Vert^2$ is the distance of an oblique projection to the kernel of $A$ which is a quasi strongly convex function. 

@@

        The above definition is the special case of relatively strong convexity wrt to the Bregman divergence $D_h(x, y) = \Vert Ax - Ay\Vert^2$. 
        \begin{theorem}[an equivalent definition for SQ-SCNVX functions]\label{thm:sq-scnvx-equiv}
            A function $F$ is semi quasi strongly convex with $A$ if and only if the function $x \mapsto F(x) - \frac{1}{2}\Vert A x\Vert^2$ is convex. 
            For all $\lambda \in \RR, x \in \RR^n, y \in \RR^n$ it has $-1/2\Vert A(\lambda x + (1 - \lambda)y)\Vert^2 = (1/2)(\lambda\Vert Ax\Vert^2 + (1 - \lambda)\Vert Ay\Vert^2 - \lambda(1 - \lambda)\Vert Ax - Ay\Vert^2)$ by verifying: 
                & -\frac{1}{2}\Vert A(\lambda x + (1 - \lambda)y)\Vert^2 + 
                    \frac{\lambda}{2}\Vert Ax\Vert^2 + \frac{1 - \lambda}{2}\Vert Ay\Vert^2 - \frac{\lambda(1 - \lambda)}{2}\Vert Ay - Ax\Vert^2
                -\frac{1}{2}\left(
                    \lambda^2\Vert Ax\Vert^2 + (1 - \lambda)^2\Vert Ay\Vert^2 - 2\lambda(1 - \lambda) \langle Ax, Ay\rangle
                        \frac{\lambda}{2} - \frac{\lambda(1 - \lambda)}{2}
                    \right)\Vert Ax\Vert^2 + \left(
                        \frac{1 - \lambda}{2} - \frac{\lambda(1 - \lambda)}{2}
                    \right)\Vert Ay\Vert^2 - \lambda(1 - \lambda)\langle Ay, Ax\rangle
                &= - \frac{\lambda^2}{2}\Vert Ax\Vert^2 - \frac{(1 - \lambda)^2}{2}\Vert Ay\Vert^2
                        \frac{\lambda}{2} - \frac{\lambda - \lambda^2}{2}
                    \right)\Vert Ax\Vert^2 + \left(
                        \frac{1 - \lambda}{2} - \frac{\lambda - \lambda^2}{2}
                    \right)\Vert Ay\Vert^2
                &= 0
            Using the above result we can prove the equivalency because 
            {\small
                0 &\le F(\lambda x + (1 - \lambda)y) + \lambda F(x) + (1 - \lambda)F(y) -\frac{\lambda(1 - \lambda)}{2} \Vert Ay - Ax\Vert^2
                &=  F(\lambda x + (1 - \lambda)y) - \frac{1}{2}\Vert A(\lambda x + (1 - \lambda)y)\Vert^2 
                + \lambda F(x) - \frac{\lambda}{2}\Vert Ax\Vert^2 
                + (1 - \lambda)F(y) - \frac{1 - \lambda}{2} \Vert Ay\Vert^2 
                    - \frac{\lambda(1 - \lambda)}{2} \Vert Ay - Ax\Vert^2 + \frac{1}{2}\Vert A(\lambda x + (1 - \lambda)y)\Vert^2
                    + \frac{1}{2}\Vert Ax\Vert^2 + \frac{1}{2}\Vert Ay\Vert^2 
                F(\lambda x + (1 - \lambda)y) - \frac{1}{2}\Vert A(\lambda x + (1 - \lambda)y)\Vert^2 
                    + \lambda \left(F(x) - \frac{1}{2}\Vert Ax\Vert^2\right) 
                    + (1 - \lambda)\left(F(y) - \frac{1}{2} \Vert Ay\Vert^2\right). 
            So, the function $F(x) - \frac{1}{2}\Vert Ax\Vert^2$ is convex. 
        The following theorem classifies a class of semi quasi strongly convex function. 
        \begin{theorem}[Affine composite differentiable SQ-SCNVX functions]\label{thm:smooth-aff-sq-scnvs-fxn}

        \end{theorem}
        \begin{proof}
            Let $h(x) = f(Ax - b)$.

        \end{proof}
        The following definition defines the concept of relative smoothness. 
        \begin{definition}[semi relative smoothness]\label{def:semi-rel-smooth}

        \end{definition}
        We build the proximal gradient inequality for the class of SQ-SCNVX functions. 
        \begin{theorem}[proximal gradient inequality]\label{thm:pg-ineq}
            Let function $F$ satisfies Assumption \ref{ass:smooth-plus-nonsmooth}, so it's $\mu \ge 0$ strongly convex. 
            For all $x\in \RR^n$, define $x^+ = T_L(x)$, then there exists a $B \ge 0$ such that $D_f(x^+, x) \le B/2 \Vert x^+ - x\Vert^2$. 
            Then, for all $z \in \RR^n$ it satisfies proximal gradient inequality at point $x$:  
            \begin{align*}
                0&\le F(z) - F(x^+) - \frac{B}{2}\Vert z - x^+\Vert^2  + \frac{B - \mu}{2}\Vert z - x\Vert^2
                \\
                &=  F(z) - F(x^+) - \langle B(x - x^+), z - x\rangle
                - \frac{\mu}{2}\Vert z - x\Vert^2
                - \frac{B}{2}\Vert x - x^+\Vert^2. 
            \end{align*}
            Since $f$ is assumed to be $L$ Lipschitz smooth, the above condition is true for all $x, y \in \RR^n$ for all $B \ge L$. 
        \end{theorem}
        \begin{remark}
            The theorem is the same as in Nesterov's book \cite[Theorem 2.2.13]{nesterov_lectures_2018}, but with the use of proximal gradient mapping and proximal gradient instead of project gradient hence making it equivalent to the theorem in Beck's book \cite[Theorem 10.16]{beck_first-order_2017}. 
            The only generalization here is parameter $B$ which made to accommodate algorithm that implements Definition \ref{def:snapg-v2} with line search routine to determine $L_k$. 
            Each of the reference books gives a proof of the theorem. 
            But for the best consistency in notations, see Theorem 2.3 in Li and Wang \cite{li_relaxed_2025}. 
        \end{remark}

@@

                \item If, in addition, the sequence $\tilde \mu_k$ satisfies for all $k \ge 1$, $\frac{\tilde \mu_k}{L_{k - 1}} < L_{k - 1}/ L_k$, then the sequence also satisfies for all $k \ge 1$: $\alpha_k < 1$. 

@@

        \begin{definition}[SNAPG-V2]\label{def:snapg-v2}
            Let $F$ satisfies Assumption \ref{ass:sum-of-many}. 
            Let $(I_k)_{k \ge 0}$ be a list of i.i.d random variables uniformly sampled from set $\{0, 1, 2, \cdots, n\}$. 
            Initialize $v_{-1} = x_{-1}, \alpha_0 = 1$. 
            Let $\tilde \mu \ge 0$ be a constant that is fixed.
            The SNAPG generates the sequence $(y_k, x_k, v_k)_{k \ge 0}$ such that for all $k \ge 0$ they satisfy: 
                & \alpha_k \in (0, 1): (L_{k - 1}/L_k)(1 - \alpha_{k})\alpha_{k - 1}^2 = \alpha_{k}\left(\alpha_{k} - \tilde \mu/L_k\right), \\
                & \tau_k = L_k(1 - \alpha_k)\left(L_k \alpha_k - \mu^{(I_k)}\right)^{-1}, \\
                & y_k = (1 + \tau_k)^{-1}v_{k - 1} + \tau_k(1 + \tau_k)^{-1}x_{k - 1}, \\
                & L_k > 0: D_f(x_k, y_k) \le L_k/2\Vert y_k - x_k\Vert^2, \\
                & x_k =  T_{L_k}(y_k | F_{I_k}), \\
                & v_k = x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
            $\tilde \mu_k, L_k$ are not necessary a random variable because they are determined by a line-search like conditions, consequently $(\alpha_k)_{k\ge 0}$, whether they are a random variable depends on the line search procedures. 
            Otherwise, all the iterates $(x_k, y_k, z_k)$ are random variable determined by $I_k$ when conditioned on all previous $I_{k - 1}, I_{k - 2}, \ldots, I_{0}$. 
            \textcolor{red}{NEW}. One may notice that $\alpha_k$ requires $L_k$ which comes before $L_k, x_k$ which are needed in advanced for $\alpha_k$. 
            This may seem off since no algorithm can know what $L_k$ to choose in advanced to determine the line search. 
            But, it is important to note that in here, we defined a sequence of conditions on the iterates $x_k, y_k, z_k$, and auxiliary sequences $\alpha_k, L_k$ which is not a definition of any algorithm. 
            It is quantifying the conditions needed for an algorithm that actually implements it.
            For the trivial case where we don't need to worry about it is when $L_k = \max_{i = 1, \ldots, n} K^{(i)}$. 
            See Chambolle, Calatroni \cite{calatroni_backtracking_2019} for an implementation of linear search with backtracking for the FISTA algorithm, it is how one would implement it in the deterministic case. 

@@

        The previous section highlighted a generic convergence results from one iteration of the algorithm, however, there are a lot of loose ends. 
        This section will deal with those.
